---
title: "PDF Document Template, Tufte Handout"
runningheader: "Tufte Handout with R Markdown"
author: "Your Name, Licence"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    includes:
      in_header: ./latex/tufte_preamble.tex # 日本語処理のために必ず指定する
    citation_package: natbib                # 参考文献一覧を使う場合
    latex_engine: xelatex                   # 他のLaTeXエンジンは使用不可
    fig_caption: yes
    keep_tex: yes
  tufte::tufte_book:
    includes:
      in_header: ./latex/tufte_preamble.tex # 日本語処理のために必ず指定する
    citation_package: natbib                # 参考文献一覧を使う場合
    latex_engine: xelatex                   # 他のLaTeXエンジンは使用不可
    fig_caption: yes
    keep_tex: yes
classoption: a4paper     # 指定しないとデフォルトではUS Letterサイズになる
always_allow_html: true  # 一種の呪文
link-citations: true     # 参考文献一覧へのリンクを有効化
bibliography: ./bib/skeleton.bib  # 参考文献一覧ファイル（ファイル末で自動生成）
linestretch: 1.1         # 行間の指定（お好みで調整を）
mainfontoptions: "Scale=MatchUppercase"
monofontoptions: "Scale=0.96"     # 0.96がベストらしい
CJKoptions: "Scale=1.0"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tufte)

# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(echo = TRUE,                 # Rコードを出力する
                      warning = FALSE,             # Rの警告を抑止する
                      fig_caption = TRUE,
                      fig.align = 'center',        # 図を中央揃えにする
                      attr.source='.numberLines'   # Rコードに行番号表示
                      )

# tufte packageでは、余白に図などを表示させる制御があるためfig.posによる
# 図表位置の固定はできません。また、見出しは節レベル（level 2）までです。
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(dev = "cairo_pdf",       # 図の出力にPDFデバイスを使う
                        tinytex.verbose = TRUE   # デバッグ用メッセージを出力
                        )
}

df_print <- function(df = NULL, n = 3L, caption = NULL, scale_down = FALSE,
                     font_size = NULL, position = "center", full_width = FALSE,
                     head_tail = FALSE, booktabs = TRUE, ...) {
  if (!is.null(df)) {
    if (knitr::is_latex_output()) {
      if (scale_down == FALSE) {
        # latex_option = c("striped", "HOLD_position")
        latex_option = c("striped")
      } else {
        # latex_option = c("striped", "HOLD_position", "scale_down")
        latex_option = c("striped", "scale_down")
      }
      
      df %>% 
        psych::headTail(top = n,
                        bottom = dplyr::if_else(head_tail == TRUE, n, 0L)) %>% 
        kableExtra::kbl(caption  = caption, booktabs = booktabs, ...) %>%
        kableExtra::kable_styling(font_size = font_size, position = position,
                                  full_width = full_width,
                                  latex_options = latex_option)
    } else {   # 出力タイプ不明の場合
      return(NULL)
    }
  }
}
```

# 本テンプレートの使い方

前提条件

* **`tidyverse`, `knitr`, `rmarkdown`, `psych`, `kableExtra`**パッケージ^[**R** 4.x推奨]
* **RStudio**^[v1.4推奨]
* **Notoフォント**（Linux）・**ヒラギノフォント**^[検証していないので推測です]（macOS）

作成手順

1. **`tufte`, `tinytex`**パッケージをインストールする
1. `tinytex::install_tinytex()`で**tinytex**をインストールする
1. `tinytex::tlmgr_install("haranoaji")`で原の味フォントをインストールする^[Windows環境のみ]
1. `latex/tufte_preamble.tex`を開いてOS環境に見合ったフォント設定を有効にする^[不要な設定は%文字でコメントアウトします]
1. 本ドキュメントをknitする
    * 必要な$\TeX$パッケージは**tinytex**が自動的にインストールする
    * もし$\TeX$のメッセージが出た場合にはログを参考に必要なパッケージをインストール[^1]する
    * 出力フォーマットを変更したい場合はYAMLの`output`指定を変更する

[^1]: `tinytex::tlmgr_install("package")`を**RStudio**のコンソールから実行すればインストールできます

注意事項

* **tinytex**以外の$\TeX / \LaTeX$を利用する場合は手動でパッケージをインストールしてください
    * **tinytex**以外の$\TeX / \LaTeX$での動作は確認していません
    * **RStudio**での$\LaTeX$エンジン指定は必ず`xelatex`を指定してください
* 本テンプレートは必要最低限の設定だけです
    * $\TeX / \LaTeX$のデフォルト仕様として図表は自動的に再配置されます^[**`tufte`**パッケージの仕様上図表の固定指定はできません]
    * 各種の指定方法は本ドキュメントに記述されています^[[ドキュメントサンプル(PDF)](https://ftp.jaist.ac.jp/pub/CTAN/macros/latex/contrib/tufte-latex/sample-handout.pdf)も参照してください]
* Winodws環境はレンダリングに時間がかかる場合があります
* 平仮名の「う(U)」が表示されない問題があります^[Linux環境で原ノ味フォントを指定した場合、Linux環境ではNotoフォントを指定してください]

\newpage

# 独自の関数定義
PDFではインタラクティブな表が使えません。また、tufteは余白が広いので通常の表出力では表示できる項目数が限られてしまいます。そこで、表現の自由度を高めるために**`kableExtra`**パッケージと**`psych`**パッケージを用いた`df_print()`関数^[詳細は`setup`チャンク内の関数が定義を参照方]を定義してあります。以下は使い方の一例です。
```{r}
mtcars[1:6, 1:6] %>% 
  df_print(caption = "デフォルトの表示方法です")

mtcars[, 1:6] %>% 
  df_print(caption = "データの先頭と最後から規定行数表示します",
           head_tail = TRUE)

mtcars %>% 
  df_print(caption = "全カラムを収めるためにスケールダウン表示します",
           scale_down = TRUE, head_tail = TRUE)
```

\newpage

# Introduction

The Tufte handout style is a style that Edward Tufte uses in his books and handouts. Tufte's style is known for its extensive use of sidenotes, tight integration of graphics with text, and well-set typography. This style has been implemented in LaTeX and HTML/CSS^[See Github repositories [tufte-latex](https://github.com/tufte-latex/tufte-latex) and [tufte-css](https://github.com/edwardtufte/tufte-css)], respectively. We have ported both implementations into the [**tufte** package](https://github.com/rstudio/tufte). If you want LaTeX/PDF output, you may use the `tufte_handout` format for handouts, and `tufte_book` for books. For HTML output, use `tufte_html`. These formats can be either specified in the YAML metadata at the beginning of an R Markdown document (see an example below), or passed to the `rmarkdown::render()` function. See @R-rmarkdown for more information about **rmarkdown**.

```yaml
---
title: "An Example Using the Tufte Style"
author: "John Smith"
output:
  tufte::tufte_handout: default
  tufte::tufte_html: default
---
```

There are two goals of this package:

1. To produce both PDF and HTML output with similar styles from the same R Markdown document;
1. To provide simple syntax to write elements of the Tufte style such as side notes and margin figures, e.g. when you want a margin figure, all you need to do is the chunk option `fig.margin = TRUE`, and we will take care of the details for you, so you never need to think about `\begin{marginfigure} \end{marginfigure}` or `<span class="marginfigure"> </span>`; the LaTeX and HTML code under the hood may be complicated, but you never need to learn or write such code.

If you have any feature requests or find bugs in **tufte**, please do not hesitate to file them to https://github.com/rstudio/tufte/issues. For general questions, you may ask them on StackOverflow: https://stackoverflow.com/tags/rmarkdown.

# Headings

This style provides first and second-level headings (that is, `#` and `##`), demonstrated in the next section. You may get unexpected output if you try to use `###` and smaller headings.

`r newthought('In his later books')`^[[Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)], Tufte starts each section with a bit of vertical space, a non-indented paragraph, and sets the first few words of the sentence in small caps. To accomplish this using this style, call the `newthought()` function in **tufte** in an _inline R expression_ `` `r ` `` as demonstrated at the beginning of this paragraph.^[Note you should not assume **tufte** has been attached to your R session. You should either `library(tufte)` in your R Markdown document before you call `newthought()`, or use `tufte::newthought()`.]

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE, message=FALSE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can adjust the proportions of figures using the `fig.width` and `fig.height` chunk options. These are specified in inches, and will be automatically scaled down to fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named `marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk starting with ```` ```{marginfigure} ```` instead, then put the content in the chunk. See an example on the right about the first fundamental theorem of calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin content as simple as possible (syntax-wise) in the `marginefigure` blocks. You may use simple Markdown syntax like `**bold**` and `_italic_` text, but please refrain from using footnotes, citations, or block-level elements (e.g. blockquotes and lists) there.

Note: if you set `echo = FALSE` in your global chunk options, you will have to add `echo = TRUE` to the chunk to display a margin figure, for example ```` ```{marginfigure, echo = TRUE} ````.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, message=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

Other chunk options related to figures can still be used, such as `fig.width`, `fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width` is large and `fig.height` is small. In the above example, the plot size is $10 \times 2$.

## Arbitrary Full Width Content

Any content can span to the full width of the page. This feature requires Pandoc 2.0 or above. All you need is to put your content in a fenced `Div` with the class `fullwidth`, e.g.,

```md
::: {.fullwidth}
Any _full width_ content here.
:::
```

Below is an example:

::: {.fullwidth}
_R is free software and comes with ABSOLUTELY NO WARRANTY._ You are welcome to redistribute it under the terms of the GNU General Public License versions 2 or 3. For more information about these matters see https://www.gnu.org/licenses/.
:::

## Main Column Figures

Besides margin and full width figures, you can of course also include figures constrained to the main column. This is the default type of figures in the LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the extensive use of sidenotes. There is a wide margin to provide ample room for sidenotes and small figures. Any use of a footnote will automatically be converted to a sidenote. ^[This is a sidenote that was entered using a footnote.] 

If you'd like to place ancillary information in the margin without the sidenote mark (the superscript number), you can use the `margin_note()` function from **tufte** in an inline R expression. `r margin_note("This is a margin note.  Notice that there is no number preceding the note.")` This function does not process the text with Pandoc, so Markdown syntax will not work here. If you need to write anything in Markdown syntax, please use the `marginfigure` block described previously.

# References

References can be displayed as margin notes for HTML output. For example, we can cite R here [@R-base]. To enable this feature, you must set `link-citations: yes` in the YAML metadata, and the version of `pandoc-citeproc` should be at least 0.7.2. You can always install your own version of Pandoc from https://pandoc.org/installing.html if the version is not sufficient. To check the version of `pandoc-citeproc` in your system, you may run this in R:

```{r eval=FALSE}
system2('pandoc-citeproc', '--version')
```

If your version of `pandoc-citeproc` is too low, or you did not set `link-citations: yes` in YAML, references in the HTML output will be placed at the end of the output document.

# Tables

You can use the `kable()` function from the **knitr** package to format tables that integrate well with the rest of the Tufte handout style. The table captions are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```

# Block Quotes

We know from the Markdown syntax that paragraphs that start with `>` are converted to block quotes. If you want to add a right-aligned footer for the quote, you may use the function `quote_footer()` from **tufte** in an inline R expression. Here is an example:

> "If it weren't for my lawyer, I'd still be in prison. It went a lot faster with two people digging."
>
> `r tufte::quote_footer('--- Joe Martin')`

Without using `quote_footer()`, it looks like this (the second line is just a normal paragraph):

> "Great people talk about ideas, average people talk about things, and small people talk about wine."
>
> --- Fran Lebowitz

# Responsiveness

The HTML page is responsive in the sense that when the page width is smaller than 760px, sidenotes and margin notes will be hidden by default. For sidenotes, you can click their numbers (the superscripts) to toggle their visibility. For margin notes, you may click the circled plus signs to toggle visibility.

# More Examples

The rest of this document consists of a few test cases to make sure everything still works well in slightly more complicated scenarios. First we generate two plots in one figure environment with the chunk option `fig.show = 'hold'`:

```{r fig-two-together, fig.cap="Two plots in one figure environment.", fig.show='hold', cache=TRUE, message=FALSE}
p <- ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point()
p
p + geom_smooth()
```

Then two plots in separate figure environments (the code is identical to the previous code chunk, but the chunk option is the default `fig.show = 'asis'` now):

```{r fig-two-separate, ref.label='fig-two-together', fig.cap=sprintf("Two plots in separate figure environments (the %s plot).", c("first", "second")), cache=TRUE, message=FALSE}
```

You may have noticed that the two figures have different captions, and that is because we used a character vector of length 2 for the chunk option `fig.cap` (something like `fig.cap = c('first plot', 'second plot')`).

Next we show multiple plots in margin figures. Similarly, two plots in the same figure environment in the margin:

```{r fig-margin-together, fig.margin=TRUE, fig.show='hold', fig.cap="Two plots in one figure environment in the margin.", fig.width=3.5, fig.height=2.5, cache=TRUE}
p
p + geom_smooth(method = 'lm')
```

Then two plots from the same code chunk placed in different figure environments:

```{r fig-margin-separate, fig.margin=TRUE, fig.cap=sprintf("Two plots in separate figure environments in the margin (the %s plot).", c("first", "second")), fig.width=3.5, fig.height=2.5, cache=TRUE}
knitr::kable(head(iris, 15))
p
knitr::kable(head(iris, 12))
p + geom_smooth(method = 'lm')
knitr::kable(head(iris, 5))
```

We blended some tables in the above code chunk only as _placeholders_ to make sure there is enough vertical space among the margin figures, otherwise they will be stacked tightly together. For a practical document, you should not insert too many margin figures consecutively and make the margin crowded. 

You do not have to assign captions to figures. We show three figures with no captions below in the margin, in the main column, and in full width, respectively.

```{r fig-nocap-margin, fig.margin=TRUE, fig.width=3.5, fig.height=2, cache=TRUE}
# a boxplot of weight vs transmission; this figure
# will be placed in the margin
ggplot(mtcars2, aes(am, wt)) + geom_boxplot() +
  coord_flip()
```
```{r fig-nocap-main, cache=TRUE}
# a figure in the main column
p <- ggplot(mtcars, aes(wt, hp)) + geom_point()
p
```
```{r fig-nocap-fullwidth, fig.fullwidth=TRUE, fig.width=10, fig.height=3, cache=TRUE}
# a fullwidth figure
p + geom_smooth(method = 'lm') + facet_grid(~ gear)
```

# Some Notes on Tufte CSS

There are a few other things in Tufte CSS that we have not mentioned so far. If you prefer `r sans_serif('sans-serif fonts')`, use the function `sans_serif()` in **tufte**. For epigraphs, you may use a pair of underscores to make the paragraph italic in a block quote, e.g.

> _I can win an argument on any topic, against any opponent. People know this, and steer clear of me at parties. Often, as a sign of their great respect, they don't even invite me._
>
> `r quote_footer('--- Dave Barry')`

We hope you will enjoy the simplicity of R Markdown and this R package, and we sincerely thank the authors of the Tufte-CSS and Tufte-LaTeX projects for developing the beautiful CSS and LaTeX classes. Our **tufte** package would not have been possible without their heavy lifting.

You can turn on/off some features of the Tufte style in HTML output. The default features enabled are:

```yaml
output:
  tufte::tufte_html:
    tufte_features: ["fonts", "background", "italics"]
```

If you do not want the page background to be lightyellow, you can remove `background` from `tufte_features`. You can also customize the style of the HTML page via a CSS file. For example, if you do not want the subtitle to be italic, you can define

```css
h3.subtitle em {
  font-style: normal;
}
```

in, say, a CSS file `my_style.css` (under the same directory of your Rmd document), and apply it to your HTML output via the `css` option, e.g.,

```yaml
output:
  tufte::tufte_html:
    tufte_features: ["fonts", "background"]
    css: "my_style.css"
```

There is also a variant of the Tufte style in HTML/CSS named "[Envisoned CSS](https://github.com/nogginfuel/envisioned-css)". This style can be used by specifying the argument `tufte_variant = 'envisioned'` in `tufte_html()`^[The actual Envisioned CSS was not used in the **tufte** package. We only changed the fonts, background color, and text color based on the default Tufte style.], e.g.

```yaml
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
```

To see the R Markdown source of this example document, you may follow [this link to Github](https://github.com/rstudio/tufte/raw/master/inst/rmarkdown/templates/tufte_html/skeleton/skeleton.Rmd), use the wizard in RStudio IDE (`File -> New File -> R Markdown -> From Template`), or open the Rmd file in the package:

```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

This document is also available in [Chinese](https://rstudio.github.io/tufte/cn/), and its `envisioned` style can be found [here](https://rstudio.github.io/tufte/envisioned/).

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = './bib/skeleton.bib')
```

